Cycles of the logistic map

### 2013-10-07 doc/pc-ijbc.tex
small corrections to the text
add a script rmcmt.py to remove comments from pc-ijbc.tex and form IJBCxxx.pdf

\documentclass{ws-ijbc}
\usepackage{multirow}
\usepackage{centernot}
%\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{calc}
\usepackage{cancel} % strike out (cancel) in math mode
\usepackage{ws-rotating}
%\usepackage{multicol}
\begin{document}

@@

%\newcommand{\eq}{}
%\newcommand{\eqs}{}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\req}[1]{(\ref{eq:#1})}
\newcommand{\refeq}[1]{\eq\,\req{#1}}
\newcommand{\refeqs}[1]{\eqs\,\req{#1}}
\newcommand{\reqsub}[2]{(\ref{eq:#1}#2)}
\newcommand{\refeqsub}[2]{\eq\,\reqsub{#1}{#2}}
\newcommand{\refeqssub}[2]{\eqs\,\reqsub{#1}{#2}}

\newcommand{\refthm}[1]{Theorem \ref{thm:#1}}
\newcommand{\refthms}[1]{Theorems \ref{thm:#1}}
\newcommand{\refsec}[1]{Section \ref{sec:#1}}
\newcommand{\refsecs}[1]{Sections \ref{sec:#1}}
\newcommand{\refapd}[1]{Appendix \ref{apd:#1}}
\newcommand{\reftab}[1]{Table \ref{tab:#1}}
\newcommand{\reftabs}[1]{Tables \ref{tab:#1}}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}
\newcommand{\reffigs}[1]{Figs. \ref{fig:#1}}

% ws-ijbc doesn't like the following lines
%\newtheorem{theorem}{Theorem}
%\newenvironment{remark}[1][1]%
%{\par\noindent\textbf{Remark #1.} }{\medskip}

@@

\title{Cycles of the logistic map}
\author{Cheng Zhang}
\address{Applied Physics Program \& Department of Bioengineering,
Rice University, Houston, TX 77005, USA}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
%The iterated sequence $x_1$,
%  $x_2 = f(x_1)$,
%  $x_3 = f(x_2)$,
%  $\ldots$,
%can be conveniently visualized on the cobweb plot,
%  as shown in \reffig{cobweb}.
%%
%Starting from $(x_1, x_1)$ on the diagonal,
%  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
%  where $y = x_{k+1} = f(x_k)$;
%by the following horizontal arrow,
%  it returns to the diagonal at
%  $(y, y) = (x_{k+1}, x_{k+1})$,
%  which starts the next iteration.
%
%
%\begin{figure}[h]
%  \begin{center}
%  \begin{minipage}{\linewidth}
%        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
%  \end{minipage}%
%  \end{center}
%  \caption{\label{fig:cobweb}
%  Cobweb plot of the logistic map.}
%\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a fixed point (possibly at the infinity),
      %e.g., \reffig{cobweb}(a);
  (ii) a periodic cycle,
      %e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic or aperiodic trajectory.
      %e.g., \reffig{cobweb}(f).
We only discuss the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
A stable fixed point requires that a small deviation from $x^*$
  gets reduced in magnitude after an iteration of \refeq{logmap}.
%
Since $f$ is differentiable,
  the condition is equivalent to $|f'(x^*)| \le 1$
  \cite{strogatz, hao}.
%
An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
By definition, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
(but the latter also permits an $x_k$ in a $d$-cycle with $d|n$).
%
%\big[but the converse is untrue, for a fixed point of $f^n$
%  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
%  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, the stability of a cycle can be defined
  according to that of the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx_1} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_n) \cdots f'(x_1) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx_1} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz, hao}.
%
%
%
The cycle and its stability depend on the parameter $r$,
%
and a stable cycle corresponds to a window
  between the onset point $r_a$ and bifurcation point $r_b$.
%
There are generally multiple such windows for a single $n$,
  but all $r_a$ are zeros of a polynomial,
  and all $r_b$ another.
%
Our aim is thus to find the two polynomials for a given $n$.
%
The problem has been tackled either by hand for small $n$
\cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
      saha, bechhoefer, gordon, burm, zhang}
or by general computer algebra methods \cite{bailey1, bailey2, kk1, lewis}.
%
Below we present a specialized algorithm to do so.
%
%
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
%Here, if $r > 0$, then
%  $r_a$ ($r_b$) are the onset (bifurcation) points.
%

@@

To simplify the calculation, we first change variables \cite{mandelbrot, brown1} as
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

@@

At first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \refeq{der}
  to get a polynomial equation of $R$.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1},
  which gradually eliminates $x_k$ out of the $n+1$ equations.
If $x_2, \dots, x_n$ are eliminated first,
  there are only two polynomial equations left:
  $f^n(x_1) - x_1 = 0$
  and
  $\frac{d}{dx_1}f^n (x_1) \mp 1 = 0$
  [\refeq{der}];
  the final elimination of $x_1$ can then be accomplished
  by setting the resultant of the two polynomials
  to zero \cite{burm}.
  %leading to the desired polynomial equation of $R$.
%
These approaches, however, have not fully
  exploited the cyclic structure of \refeqs{logmaps},
  and can be improved by the following alternative.

Instead of solving \refeqs{logmaps} for $x_k$,
  we will derive a set of equations
  in terms of the cyclic polynomials $\{C_p\}$ of $x_k$.
  %(an example of a cyclic polynomial is
  %$x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
  %with $n$ terms).
%
The advantage is that the new set of equations are
  \emph{homogeneous}, \emph{linear} and \emph{complete},
  i.e., there is an equation
  $\sum_q A_{pq}(R) C_q = 0$ for every cyclic polynomial $C_p$.
%
The matrix $\mathbf A(R)$ formed by the coefficients $A_{pq}(R)$
  must have a zero determinant,
  for the cyclic polynomials $\{C_p\}$ are not zeros altogether.
%
Thus, the zero-determinant condition $|\mathbf A(R)| = 0$ gives
  the required polynomial equation of $R$.
%whose roots contain all fixed points of $f^n$.
%

@@

A cyclic polynomial is an invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b = x_1 x_3 + x_2 x_4$ for $n = 4$.
%
A cyclic polynomial is not necessarily symmetric,
  that is, invariant under the exchange of any two $x_i$ and $x_{j}$,
  e.g., the above $a$ and $b$ are not symmetric, but $a + b$ is.

We call a cyclic polynomial simple
  if it is equal to the sum of
  all distinct cyclic versions of
  a unit-coefficient monomial of $x_k$.
%
The monomial is called a generator of the cyclic polynomial.
%
For example, the above $a$ and $b$ are simple,
  and their generators are $x_1 x_2$ and $x_1 x_3$, respectively.
%
A simple cyclic polynomial
  $C_p(\vx{n})$, with $\vx{n} \equiv \{x_1, \ldots, x_n\}$,
can be labeled as follows.
%
If the monomial generator is
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  we define the label $p$ as the sequence of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
%
For example,
$C_{12}(\vx{n})  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
and
$C_{112}(\vx{n}) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
%
Below, we may drop %the subscript ``$_n$'' or even the argument
  ``$(\vx{n})$'' when it is clear.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest label $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ over $x_2 x_3$ as the generator of $C_{12}(\vx{n})$,
  because ``$12$'' $<$ ``$23$.''
%
Finally, we add $C_0(\vx{n}) \equiv 1$.
%
%
%

@@

A cyclic polynomial is square-free
  if it has no square or higher powers of $x_k$,
e.g., $C_{12} = x_1 x_2 + \cdots$ is square-free,
  $C_{112} = {x_1}^2 x_2 + \cdots$ is not.
%see \reftab{sqrfreepoly} for more examples.
The label $p$ of a square-free polynomial
  has no repeated index.
We denote the set of all square-free $p$ by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\cdots n\}$,
and its size by $\NB$ [cf. \refeq{necklace} below for the formula].
% is the number of square-free cyclic polynomials.
%
%
The set of the simple square-free cyclic polynomials
  furnishes a basis for expanding cyclic polynomials:

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx{n})$
  formed by the $n$-cycle points
  $\vx{n} = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the simple square-free cyclic polynomials $C_p(\vx{n})$:
\[
  K(\vx{n}) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx{n}),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx{n})$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since both the replacement rules and the original polynomial are cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx{n}) \equiv 1$.
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only new variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vx{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
We will further show below that
  at the onset and bifurcation points,
  the $\NB$ square-free cyclic polynomials $C_p$ are themselves
  linearly connected by a characteristic equation
  furnished by \refeq{der},
  which is the key step of the solution.

@@

First, the derivative $\frac{d}{dx_1}f^n(x_1)$ is
  a cyclic polynomial of $\vx{n} = \{x_1, \ldots, x_n\}$:
%
%
\begin{equation}
  \Lambda(\vx{n})
   = f'(x_n) \dots f'(x_1)
   = (-2)^n x_1 \dots x_n.
\label{eq:logder}
\end{equation}
%
%
For any square-free cyclic polynomial $C_p(\vx{n})$,
the product $\Lambda(\vx{n}) \, C_p(\vx{n})$ is also cyclic,
  and can thus be expanded by \refthm{sqrfree} as
\begin{equation}
  \Lambda(\vx{n}) \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n}),
\label{eq:xcp0}
\end{equation}
where $T_{pq}(R)$ is a polynomial of $R$,
and $p, q \in \Pset$.

@@

Since a set of homogeneous linear equations
  has a nonzero solution of $\vct C$
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\Rlam \equiv \big| \vct T(R) - \lambda \, \vct I \big| = 0.
\label{eq:xr}
\end{equation}
%
This determinant $A_n\Rlam$
  is the desired polynomial of $R$ and $\lambda$.
%
%The attached subscript $n$ allows us
%  to use the determinant later with
%  similar determinants of divisors $d$ of $n$.
%
To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx{n})$
  of the cycle points $\vx{n} = \{x_k\}$
  are linearly related as
  $\lambda \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n})$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx{n}) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed points of $f^n$
  are marginally stable when $R$ satisfies
  the characteristic equation
  $\big| \vct T(R) - \lambda \, \vct I \big| = 0$.
  \label{thm:main}
\end{theorem}

@@

We illustrate the above algorithm on several small-$n$ cases.
%
%It is still helpful to have a mathematical software
%  at hand to verify certain steps
%  (e.g., in computing the determinants and their factorization).
%
%
For $n = 1$, % (the fixed point),
we have two simple square-free cyclic polynomials
$\{C_0, C_1\} = \{1, x_1\}$, with
$\Lambda = -2 \, x_1$. So
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
The characteristic equation (\ref{eq:xr}) reads
\begin{equation}
A_1\Rlam
  %=\left| \begin{array}{cc}
  %\lambda     & 2          \\
  %2 R         & \lambda - 2
  %\end{array} \right|
  \equiv -4R - 2\lambda + \lambda^2
  = 0.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
So $R = (\lambda^2 -2\lambda)/4$,
  so the fixed point becomes unstable at $R_b = 3/4$ where $\lambda = -1$.

>>>>

\documentclass{ws-ijbc}
\usepackage{multirow}
%\usepackage{centernot}
%\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage{calc}
\usepackage{cancel} % strike out (cancel) in math mode
\usepackage{ws-rotating}
%\usepackage{multicol}
\begin{document}

@@

\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\req}[1]{(\ref{eq:#1})}
\newcommand{\refeq}[1]{\eq\,\req{#1}}
\newcommand{\refeqs}[1]{\eqs\,\req{#1}}
\newcommand{\reqsub}[2]{(\ref{eq:#1}#2)}
\newcommand{\refeqsub}[2]{\eq\,\reqsub{#1}{#2}}
\newcommand{\refeqssub}[2]{\eqs\,\reqsub{#1}{#2}}

\newcommand{\refthm}[1]{Theorem \ref{thm:#1}}
\newcommand{\reflem}[1]{Lemma \ref{thm:#1}}
\newcommand{\refsec}[1]{Section \ref{sec:#1}}
\newcommand{\refsecs}[1]{Sections \ref{sec:#1}}
\newcommand{\reffig}[1]{Fig. \ref{fig:#1}}

@@

\title{Cycles of the logistic map}
\author{Cheng Zhang}
\address{Applied Physics Program and Department of Bioengineering,
Rice University, Houston, TX 77005, USA}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
%The iterated sequence $x_1$,
%  $x_2 = f(x_1)$,
%  $x_3 = f(x_2)$,
%  $\ldots$,
%can be conveniently visualized on the cobweb plot,
%%
%Starting from $(x_1, x_1)$ on the diagonal,
%  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
%  where $y = x_{k+1} = f(x_k)$;
%by the following horizontal arrow,
%  it returns to the diagonal at
%  $(y, y) = (x_{k+1}, x_{k+1})$,
%  which starts the next iteration.
%
%
%\begin{figure}[h]
%  \begin{center}
%  \begin{minipage}{\linewidth}
%        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
%  \end{minipage}%
%  \end{center}
%  \caption{\label{fig:cobweb}
%  Cobweb plot of the logistic map.}
%\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a fixed point (possibly at the infinity),
      %e.g., \reffig{cobweb}(a);
  (ii) a periodic cycle,
      %e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic or aperiodic trajectory.
      %e.g., \reffig{cobweb}(f).
We only discuss the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
A stable fixed point requires that a small deviation from $x^*$
  gets reduced in magnitude after an iteration of \refeq{logmap}.
%
Since $f$ is differentiable,
  the condition means $|f'(x^*)| \le 1$
  \cite{strogatz, hao}.
%
An $n$-cycle is a sequence $x_1, \dots, x_n$
  such that $n$ is the smallest positive integer
  satisfying $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
By definition, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
(whereas the latter also permits an $x_k$ in any $d$-cycle with $d|n$).
%
%\big[but the converse is untrue, for a fixed point of $f^n$
%  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
%  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, the stability of a cycle can be defined
  according to that of the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx_1} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_n) \cdots f'(x_1) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx_1} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz, hao}.
%
%
%
The cycle and its stability depend on the parameter $r$,
%
and a stable cycle corresponds to a window of $r$
  between the onset point $r_a$ and bifurcation point $r_b$.
%
There are generally multiple such windows for a single $n$,
  but all $r_a$ are zeros of a polynomial,
  and all $r_b$ zeros of another.
%
Our aim is thus to find the two polynomials for a given $n$.
%
The problem was tackled either by hand for small $n$
\cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
      saha, bechhoefer, gordon, burm, zhang}
or by general computer algebra methods \cite{bailey1, bailey2, kk1, lewis}.
%
Below we present a specialized algorithm, which offers some performance gain.
%
%
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
%Here, if $r > 0$, then
%  $r_a$ ($r_b$) are the onset (bifurcation) points.
%

@@

To simplify the calculation, we first change variables \cite{mandelbrot, brown1} as
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as the zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

@@

At first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \refeq{der}
  to get an equation of $R$.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1},
  which gradually eliminates $x_k$ out of the $n+1$ equations.
If $x_2, \dots, x_n$ are eliminated first,
  there are only two polynomial equations left:
  $f^n(x_1) - x_1 = 0$
  and
  $\frac{d}{dx_1}f^n (x_1) \mp 1 = 0$
  [\refeq{der}];
  the final elimination of $x_1$ can then be accomplished
  by setting the resultant of the two polynomials
  to zero \cite{burm}.
  %leading to the desired polynomial equation of $R$.
%
These approaches, however, have not fully
  exploited the cyclic structure of \refeqs{logmaps},
  and can be improved by the following alternative.

Instead of solving \refeqs{logmaps} for $x_k$,
  we will derive a set of equations
  in terms of the cyclic polynomials $\{C_p\}$ of $x_k$.
  %(an example of a cyclic polynomial is
  %$x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
  %with $n$ terms).
%
The advantage is that the new set of equations are
  \emph{homogeneous}, \emph{linear} and \emph{complete},
  i.e., there is an equation
  $\sum_q A_{pq}(R) C_q = 0$ for every cyclic polynomial $C_p$.
%
The square matrix $\mathbf A(R)$ formed by the coefficients $A_{pq}(R)$
  must have a zero determinant,
  for the cyclic polynomials $\{C_p\}$ are not zeros altogether.
%
Thus, the zero-determinant condition $|\mathbf A(R)| = 0$ gives
  the required polynomial equation of $R$.
%whose roots contain all fixed points of $f^n$.
%

@@

A cyclic polynomial is an invariant
  under the cycling of the variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b = x_1 x_3 + x_2 x_4$ for $n = 4$.
%
A cyclic polynomial is not necessarily symmetric,
  that is, invariant under the exchange of any two $x_i$ and $x_{j}$,
  e.g., the above $a$ and $b$ are not symmetric, but $a + b$ is.

We call a cyclic polynomial simple
  if it is equal to the sum of
  all distinct cyclic versions of
  a unit-coefficient monomial of $x_k$.
%
The monomial is called a generator of the cyclic polynomial.
%
For example, the above $a$ and $b$ are simple,
  and $x_1 x_2$ and $x_1 x_3$ are their generators, respectively.
%
A simple cyclic polynomial
  $C_p(\vx{n})$, with $\vx{n} \equiv \{x_1, \ldots, x_n\}$,
can be labeled as follows.
%
If the monomial generator is
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  we define the label $p$ as the sequence of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s.
%
For example,
$C_{12}(\vx{n})  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
and
$C_{112}(\vx{n}) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
%
Below we may drop %the subscript ``$_n$'' or even the argument
  ``$(\vx{n})$'' when $n$ is clear from the context.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest label $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ over $x_2 x_3$ as the generator of $C_{12}(\vx{n})$,
  because ``$12$'' $<$ ``$23$.''
%
Finally, we add $C_0(\vx{n}) \equiv 1$.
%
%
%

@@

A cyclic polynomial is square-free
  if it has no square or higher powers of $x_k$,
e.g., $C_{12} = x_1 x_2 + \cdots$ is square-free,
  $C_{112} = {x_1}^2 x_2 + \cdots$ is not.
The label $p$ of a simple square-free cyclic polynomial
  has no repeated index.
We denote the set of labels $p$ of all simple square-free cyclic polynomials by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\cdots n\}$,
and its size by $\NB$ [cf. \refeq{necklace} below for the formula].
% is the number of square-free cyclic polynomials.
%
%
The set of the simple square-free cyclic polynomials
  furnishes a basis for expanding an arbitrary cyclic polynomial:

\begin{lemma}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx{n})$
  formed by the $n$-cycle points
  $\vx{n} = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the simple square-free cyclic polynomials $C_p(\vx{n})$:
\[
  K(\vx{n}) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx{n}),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of labels of all simple square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{lemma}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx{n})$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since both the replacement and the original polynomial are cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx{n}) \equiv 1$.
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only new variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vx{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\reflem{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the simple square-free ones.
The key step of the solution is to further show that
  at the onset and bifurcation points,
  the $\NB$ simple square-free cyclic polynomials $C_p$ are themselves
  linearly connected by a characteristic equation
  furnished by \refeq{der}.

@@

First, the derivative $\frac{d}{dx_1}f^n(x_1)$ is
  a cyclic polynomial of $\vx{n} = \{x_1, \ldots, x_n\}$:
%
%
\begin{equation}
  \Lambda(\vx{n})
   = f'(x_n) \dots f'(x_1)
   = (-2)^n x_1 \dots x_n.
\label{eq:logder}
\end{equation}
%
%
For any simple square-free cyclic polynomial $C_p(\vx{n})$,
the product $\Lambda(\vx{n}) \, C_p(\vx{n})$ is also cyclic,
  and can thus be expanded by \reflem{sqrfree} as
\begin{equation}
  \Lambda(\vx{n}) \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n}),
\label{eq:xcp0}
\end{equation}
where $T_{pq}(R)$ is a polynomial of $R$,
and $p, q \in \Pset$.

@@

Since the set of the homogeneous linear equations
  has a nonzero solution of $\vct C$
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\Rlam \equiv \big| \vct T(R) - \lambda \, \vct I \big| = 0.
\label{eq:xr}
\end{equation}
%
This determinant $A_n\Rlam$
  is the desired polynomial of $R$ and $\lambda$.
%
%The attached subscript $n$ allows us
%  to use the determinant later with
%  similar determinants of divisors $d$ of $n$.
%
To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the simple square-free cyclic polynomials $C_p(\vx{n})$
  of the cycle points $\vx{n} = \{x_k\}$
  are linearly related as
  $\lambda \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n})$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx{n}) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed points of $f^n$
  are marginally stable when $R$ satisfies
  the characteristic equation
  $\big| \vct T(R) - \lambda \, \vct I \big| = 0$.
  \label{thm:main}
\end{theorem}

@@

We illustrate the above algorithm on several small-$n$ cases.
%
%It is still helpful to have a mathematical software
%  at hand to verify certain steps
%  (e.g., in computing the determinants and their factorization).
%
%
For $n = 1$, % (the fixed point),
we have only two simple square-free cyclic polynomials
$\{C_0, C_1\} = \{1, x_1\}$, with
$\Lambda = -2 \, x_1$. So
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
The characteristic equation (\ref{eq:xr}) reads
\begin{equation}
A_1\Rlam
  %=\left| \begin{array}{cc}
  %\lambda     & 2          \\
  %2 R         & \lambda - 2
  %\end{array} \right|
  \equiv -4R - 2\lambda + \lambda^2
  = 0.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
So $R = (\lambda^2 -2\lambda)/4$,
  and the fixed point becomes unstable at $R_b = 3/4$ where $\lambda = -1$.
